### 2015-07-11 background-new.tex
intro and background revision: incorporated comments until background 2.1

In comparison, following are the main new research directions that we pursue in this thesis. We explain each of them further in the chapter.

\begin{itemize}
	\item
	How do ISP traffic engineering schemes compare in terms of application-performance metrics such as TCP download time? Further, how do they compare while accounting fo r the effect of  application adaptation? 
	\item
	How should network CDNs--an ISP that operates a CDN on its infrastructure to deliver content to users on its network--perform content delivery given that they control both the content delivery and the underlying network routing?
	\item
	What role does content placement influence the interaction between network and content delivery, and what its its effect on user-perceived performance and netowrk cost?
\end{itemize}

@@

\begin{itemize}
	\item
	ISP traffic engineering is studied mostly as an optimization problem for minimizing link-utilization based cost functions (Section \ref{sec:ch2-te}).
	\item
	Content delivery networks (CDNs) commonly use \emph{demand-oblivious} techniques for content placement and request redirection towards improving user-perceived performance across the Internet (Section \ref{sec:ch2-cdn}).
	\item
	Interaction between traffic engineering and content delivery has been shown to affects user-perceived performance as well as network cost. But, prior work has focused mostly on the interaction of traffic engineering with either overlay routing or with request redirection (Section \ref{sec:ch2-te-cdn}).
\end{itemize}

@@

The goal of traffic engineering (TE)  is to avoid congestion hotspots in the network by optimizing routes based on network topology and expected traffic demand. In the context of large Internet service provider (ISP) networks, traffic engineering decides both intra-domain (within the ISP) and inter-domain routing (across ISPs). We focus here on intra-domain routing and refer the reader to  \cite{Feamster2003,rexford} for a survey of inter-domain traffic engineering.

@@

We classify traffic engineering approaches into three groups:
\begin{description}
	\item[Optimal] If the traffic matrix is known accurately, the optimal solution to the traffic engineering problem can formulated as a multi-commodity flow optimization problem. The routing thus computed is reffered to as \textbf{Optimal} traffic engineering in literature.
	\item[Demand-aware] As it is impossible to have accurate knowledge of future traffic matrices, a demand-aware  traffic engineering  periodically updates routing using historically observed traffic matrices.
	\item[Demand-oblivious] A demand-oblivious  traffic engineering  requires no explicit measurement of traffic matices, but instead configures routing statically. A demand-oblivious  traffic engineering  is simpler to implement because it requires neither measuremnt nor periodic updating of network routing.

@@

Content delivery systems seek to provide a high-quality experience to users accessing content in all regions  at all times. A canonical example of a content delivery system is a content delivery network (CDN). State-of-the-art CDNs operate geo-distributed datacenters, and use a combination of edge caching, intelligent server selection, and path and protocol optimizations for delivery of several types of content, e.g., video, bulk downloads, and interactive websites \cite{DilleyMPPSW02,akamai-overview}. Given their geo-distributed deployment, the decisions of content placement, i.e., locations at which a content is placed, and request redirection, i.e., which location is best positioned to serve a user's request, are central to the functioning of a CDN.

@@

\textbf{Request redirection:} Request redirection strategies complement placement strategies by selecting the server location that is best suited to process a user's request. These strategies have been extensively studied and form the heart of CDN technology today. To quote from a report by Akamai,  \emph{``the system directs client requests to the nearest available server likely to have the requested content."} where the ``nearest" server is one whose round trip latency as well as packet losses are small, and  an ``available" server is one that is lightly loaded considering all resources, i.e., network, CPU and disk  \cite{DilleyMPPSW02}.

@@

Similar to our classification of traffic engineering schemes, we consider common content delivery techniques discussed above to be demand-oblivious since they do not require explicit measurement of content demand. In contrast, a demand-aware approach to content placement and redirection has also been studied, e.g., Applegate \emph{et al.} use a demand-aware approach to determine placement and redirection for Video-on-Demand content in the network.

@@

Content delivery routes traffic at the application layer using an overlay network of servers, and traffic enginering configures network routing of the underlying physical paths. Studying this interaction between decisions at the overlay and the underlying network has been a topic of long standing interest in computer science. Several related questions have been put forth. Does the interaction between network routing and content delivery yield sub-optimal results? How can we design cooperative mechanisms to leverage these interactions? Do cooperative mechanisms yield benefit in an Internet-like environment? We find that nearly all previous research on this topic has focused either on the interaction of overlay routing with the network routing \cite{Roughgarden,selfishQiu}, or that of server selection strategies with the underlying routing  \cite{Jiang2009,JohariGameTheory, CATE, P4P} as discussed below.

@@

\textbf{Interaction between traffic engineering and overlay routing:} Several results show the negative interaction between selfish overlay routing and network routing \cite{Roughgarden,selfishQiu}. Theoretical results indicate that the negative interaction could cause an arbitrary degradation in user perceived delay. Further studies using synthetic traffic demands and topologies indicate that this interaction hurt traffic engineering metrics. However, it appears that a small fraction of Internet traffic uses overlay routing.  For example, traffic from CDN edge server to the client always follows network routing. Further, overlay routing yields ``marginal" benefits ($<$ 30\%) over network routing for 79\%-96\% of paths depending on which geographic region is being considered  \cite{rahul2006overlays}, which suggests that traffic between CDN servers forming an overlay network follows network routing in most cases. For this reason, this thesis does not model the interaction between overlay and network routing.

\textbf{Interaction between traffic engineering and request redirection:} Recent research has investigated the interaction between request redirection and traffic engineering, without considering the role of placement strategies. This interaction is commonly studied in the context of Internet service providers (ISPs) and content providers (CPs) with geo-distributed datacenters. 
Both analytical results \cite{Jiang2009,JohariGameTheory} and system implementations \cite{CATE,P4P} have shown that there is value for joint optimization of request redirection and traffic engineering, and cooperative strategies can help traffic engineering metrics and also reduce user-perceived latencies. Commonly, these efforts assume that all content is available at all locations, ignoring the fact that content availability at a location depend on placement strategies. Therefore, in this thesis, we account for the effect of content placement along with request redirection, in studying the interaction between network and content delivery.

>>>>

%In comparison, following are the main new research directions that we pursue in this thesis. We explain each of them further in the chapter.
%
%\begin{itemize}
%	\item
%	How do ISP traffic engineering schemes compare in terms of application-performance metrics such as TCP download time? Further, how do they compare while accounting fo r the effect of  application adaptation? 
%	\item
%	How should network CDNs--an ISP that operates a CDN on its infrastructure to deliver content to users on its network--perform content delivery given that they control both the content delivery and the underlying network routing?
%	\item
%	What role does content placement influence the interaction between network and content delivery, and what its its effect on user-perceived performance and netowrk cost?
%\end{itemize}

@@

\begin{itemize}
	\item ISP traffic engineering schemes commonly use a demand-aware approach, i.e., they compute routes using previously measured traffic matrics, for optimizing  link-utilization based cost functions (Section \ref{sec:ch2-te}).	\item CDNs commonly use \emph{demand-oblivious} techniques, i.e., techniques that require no system-wide measurement of content demand, for content placement and request redirection towards improving user-perceived performance across the Internet (Section \ref{sec:ch2-cdn}).
	\item Interaction between traffic engineering and content delivery affects user-perceived performance as well as network cost. But, prior work has focused mostly on the interaction of traffic engineering with two aspects of content delivery --  overlay routing and request redirection (Section \ref{sec:ch2-te-cdn}).
\end{itemize}

@@

The goal of traffic engineering is to avoid congestion hotspots in the network by optimizing routes based on network topology and expected traffic demand. In ISP networks, traffic engineering decides both intra-domain (within the ISP) and inter-domain routing (across ISPs). We focus here on intra-domain routing and refer the reader to  \cite{Feamster2003,rexford} for a survey of inter-domain traffic engineering.

@@

We classify traffic engineering approaches into three groups:
\begin{description}
	\item[Optimal] If the traffic matrix is known accurately, the optimal solution to the traffic engineering problem can formulated as a multi-commodity flow optimization problem. The routing thus computed is reffered to as Optimal traffic engineering in literature.
	\item[Demand-aware] As it is impossible to have accurate knowledge of future traffic matrices, a demand-aware  traffic engineering  periodically updates routing using historically observed traffic matrices.
	\item[Demand-oblivious] A demand-oblivious  traffic engineering  requires no explicit measurement of traffic matices, but instead configures routing statically. A demand-oblivious  traffic engineering  is simpler to implement because it requires neither measuremnt nor periodic updating of network routing.

@@

Content delivery systems seek to provide a high-quality experience to users accessing content in all regions  at all times. A canonical example of a content delivery system is a content delivery network (CDN). State-of-the-art CDNs operate geo-distributed datacenters, and use a combination of edge caching, intelligent request redirection, and path and protocol optimizations for delivery of several types of content, e.g., video, bulk downloads, and interactive websites \cite{DilleyMPPSW02,akamai-overview}. Given their geo-distributed deployment, the decisions of content placement, i.e., locations at which a content is placed, and request redirection, i.e., which location is best positioned to serve a user's request, are central to the functioning of a CDN.

@@

\textbf{Request redirection:} In a CDN, request rediection occurs at two tiers - inter-datacenter and intra-datacenter. Our focus in this part of the thesis is on inter-datacenter request redirection because it influences wide-area traffic patterns that affect traffic engineering. We discuss intra datacenter redirection in the context of a Content Datcenter (Chapter \ref{ch:shrink}).

Request redirection strategies complement placement strategies by selecting the server location that is best suited to process a user's request. These strategies have been extensively studied and form the heart of CDN technology today. To quote from a report by Akamai,  \emph{``the system directs client requests to the nearest available server likely to have the requested content."} where the ``nearest" server is one whose round trip latency as well as packet losses are small, and  an ``available" server is one that has sufficient resources to serve a request \cite{DilleyMPPSW02}.

@@

Content delivery techniques can be classified into demand-aware and demand-oblivious similar to our classfication of traffic engineering schemes. We consider common content delivery techniques discussed above to be demand-oblivious since they do not require system-wide measurement of content demand. In contrast, a demand-aware approach to content placement and redirection has also been studied, e.g., Applegate \emph{et al.} use a demand-aware approach to determine placement and redirection for Video-on-Demand content in the network.

@@

Content delivery routes traffic at the application layer using an overlay network of servers, and traffic enginering configures network routing of the underlying physical paths. Studying this interaction between decisions at the overlay and the underlying network has been a topic of long standing interest in computer science. Several related questions have been put forth. Does the interaction between network routing and content delivery yield sub-optimal results? How can we design cooperative mechanisms to leverage these interactions and do cooperative yield benefit in an Internet-like environment? We find that nearly all previous research on this topic has focused on the interaction of traffic engineering with either overlay routing \cite{Roughgarden,selfishQiu} with request redirection \cite{Jiang2009,JohariGameTheory, CATE, P4P} as discussed below.

@@

\textbf{Interaction between traffic engineering and overlay routing:} Several results show the negative interaction between selfish overlay routing and network routing \cite{Roughgarden,selfishQiu}. Theoretical results indicate that the negative interaction could cause an arbitrary degradation in user perceived delay. Further studies using ``Internet-like'' traffic demands and topologies indicate that this interaction hurts traffic engineering metrics. In this thesis, we do not consider overlay routing since our focus is primarily on intra-domain routing while overlay routing is most commonly used to mask inter-domain routing inefficiencies.

%However, it appears that a small fraction of CDN traffic uses overlay routing.  For example, traffic from CDN edge server to the client always follows network routing. Further, overlay routing yields ``marginal" benefits ($<$ 20\%) over network routing for 79\%-96\% of paths depending on which geographic region is being considered  \cite{rahul2006overlays}, which suggests that traffic between CDN servers forming an overlay network follows network routing in most cases. For this reason, this thesis does not model the interaction between overlay and network routing \tbd{is this the best reason?}.

\textbf{Interaction between traffic engineering and request redirection:} Recent research has investigated the interaction between request redirection and traffic engineering, without considering the role of placement strategies. This interaction is commonly studied in the context of Internet service providers (ISPs) and content providers (CPs) with geo-distributed datacenters. Both analytical results \cite{Jiang2009,JohariGameTheory} and system implementations \cite{CATE,P4P} have shown that there is value for joint optimization of request redirection and traffic engineering, and cooperative strategies can help traffic engineering metrics and either maintain or improve user-perceived latencies.

%Compared to these efforts, our work additionally considers content placement as a degree of freedom in a Network CDN.
%
%However, as these efforts do not model content placement as a part of the formulation it is 
%Commonly, these efforts assume that all content is available at all locations, ignoring the fact that content availability at a location depend on placement strategies. Therefore, in this thesis, we account for the effect of content placement along with request redirection, in studying the interaction between network and content delivery.
